---
title: "Semi-Supervised STACAS demo"
author: "A. Berenstein, C.Amezcua"
date: "22/07/2022"
output:
  rmdformats::readthedown:
    self-contained: true
    highlight: haddock
    thumbnails: false
    css: styles.css
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file, encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'PanBcellDef.html'))})
---

# Introduction 
This notebook will work on data produced by Amezcua et al. (Cell 2019)
In that work, Amezcua analyzed tissue resident Th cells. 
Now, the idea is to recycle those data but redirecting the focus on Bcells. In particular, those tissue resident ones. 


```{r}
renv::activate()
renv::restore()
#renv::snapshot()
#install.packages("Seurat")
#install.packages("remotes")
#remotes::install_github("carmonalab/UCell")
#remotes::install_github("carmonalab/scGate")
#install.packages("BiocManager")
#BiocManager::install("SingleR")
#BiocManager::install("celldex")
#BiocManager::install("ontoProc")
#BiocManager::install("rmarkdown")

library("Seurat")
library("dplyr")
library("UCell")
library("scGate")
library("SingleR")
library("ggplot2")
source("./utils.R")

```

# Load datasets from counts
```{r}
path <- c("~/Dropbox/Amezcua/data")
samples <- c("DS-017-WT","DS-017-Im1","DS-017-Im2")
count_files <- paste(path,samples,"gene_exon_tagged_dge.txt.gz",sep = "/")

# Output files:
qmin = 0.2
qmax = 0.9
n_var_feat = 800
res_folder = paste0("./results/","resQ_",qmin,"_",qmax,"_Nfeat",n_var_feat)
dataset_folder = paste0(res_folder,"/datasets")
plot_folder = paste0(res_folder,"/plots")
markers_folder = paste0(res_folder,"/markers")
panBcellFile <-  paste0(dataset_folder,"/seurat_processed.rds")  # Subseted Pan Bcell FIle with recomputed UMAP


# create ouput folders
dir.create(dataset_folder,recursive = T)
dir.create(plot_folder,recursive = T)
dir.create(markers_folder,recursive = T)

```

## Load proccessed object
```{r}
if(file.exists(panBcellFile)){
  run =F
  seurat.merged.processed <- seurat.pan.Bcells <- readRDS(panBellFile)  
}else{
  run =T
}

```


#create Seurat objects
```{r}
if(run){
  seuratDataList <- list()
  for(i in seq_along(samples)){
    countdata <- read.table(
      file = count_files[i],
      header = T,
      row.names = 1
    )
      
    obj <- Seurat::CreateSeuratObject(counts = countdata,project = samples[i])
    seuratDataList[[i]] <- obj
  }
  ncell <- table(unlist(sapply(seuratDataList,colnames)))
  to.drop <- names(ncell[ncell>1])
  
  # remove 4 barcodes duplicated across samples
  seuratDataList = lapply(seuratDataList,function(x){
    res = subset(x,cells = setdiff(colnames(x),to.drop) )
    return(res)
  })
}
```


# quality filtering
```{r,fig.height= 8,fig.width= 13}
if(run){
  seurat.merged <- Reduce(merge, seuratDataList)
  seurat.merged@meta.data%>%head()
  
  percent.ribo.dv <- PercentageFeatureSet(seurat.merged, pattern = "^Rp[ls]")
  percent.mito.dv <- PercentageFeatureSet(seurat.merged, pattern = "^mt-")
  
  seurat.merged <- AddMetaData(seurat.merged, metadata = percent.ribo.dv, col.name = "percent.ribo")
  seurat.merged <- AddMetaData(seurat.merged, metadata = percent.mito.dv, col.name = "percent.mito")
Seurat::VlnPlot(seurat.merged,features = c("nCount_RNA","nFeature_RNA","percent.mito", "percent.ribo"),group.by = "orig.ident",ncol = 2)
}  
```

```{r}
splited.by.sample <- SplitObject(seurat.merged, split.by = "orig.ident")
x2 <- splited.by.sample[[2]]
x3 <- splited.by.sample[[3]]
count2 <- quantile(x2@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.2,0.8,0.9,0.95,0.98,0.99))
feature2 <- quantile(x2@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.2,0.8,0.9,0.95,0.98,0.99))

count3 <- quantile(x3@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.2,0.8,0.9,0.95,0.98,0.99))
feature3 <- quantile(x3@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.2,0.8,0.9,0.95,0.98,0.99))

count_and_features = cbind(count2,count3,feature2,feature3)
write.table(count_and_features,"./results/count_and_features.tsv")
```

# Filter low or high expressed cells 
```{r}
if(run){
  splited.by.sample <- SplitObject(seurat.merged, split.by = "orig.ident")
  
  splited.by.sample.qc <- lapply(splited.by.sample, function(x) {
     #n UMIs
     q <- quantile(x@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.2,0.9,0.95,0.98,0.99))
     x <- subset(x, subset=nCount_RNA>max(q["20%"]) & nCount_RNA<q['90%']) 
     
     #n genes
     q <- quantile(x@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.2,0.9,0.95,0.98,0.99))
     x <- subset(x, subset=nFeature_RNA>max(q['20%']) & nFeature_RNA < q['90%'])
     
     #% ribo
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
     x <- subset(x, subset=percent.ribo>q['1%'] & percent.ribo < min(50,q['99%']) )
     
     #% mito
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
     x <- subset(x, subset=percent.mito < q['99%'])
     return(x)
  })
  
  seurat.filtered.merged <- Reduce(merge, splited.by.sample.qc)
  dim(seurat.merged)
  dim(seurat.filtered.merged)
}else{
  seurat.filtered.merged <- seurat.merged.processed
}

summary = sapply(c("nCount_RNA","nFeature_RNA","percent.mito", "percent.ribo"),function(x){summary(seurat.filtered.merged[[x]])})
write.table(summary,paste0(res_folder,"/summary_statistics.tsv",quote =F))
seurat.filtered.merged
```

## Control step: how many cells were dropped by high expression values? (Re: 262 cells)
```{r}
if(run){
  high.expr.dropped <- lapply(splited.by.sample, function(x){
     q <- quantile(x@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99))
     x_ncount <- subset(x, nCount_RNA>q['95%']) 
     
     #n genes
     q <- quantile(x@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99))
     x_nfeature <- subset(x, subset= nFeature_RNA > q['95%'])
     
     #% ribo
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99),na.rm =T)
     x_ribo <- subset(x, subset=percent.ribo > min(50,q['99%']))
     
     #% mito
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99),na.rm =T)
     x_mito <- subset(x, subset=percent.mito > q['99%'])
     
     return(table(c(x_ncount%>%colnames(), x_nfeature%>%colnames(),x_ribo%>%colnames(),x_ribo%>%colnames())))
    }
  )
  
  # 262 cells dropped by staying over the top 1% quantile in any of control quantity
  cells.dropped.by.hihg.quantile <- unlist(high.expr.dropped)
  cells.dropped.by.hihg.quantile%>%length()
}
```


## Sanity check: visual detection of possible bach effects (Re: thats ok)
# Process object
```{r,fig.height= 4.5,fig.width= 12}
ndim = 20      # for pca
nfeat = 1000    # number of most variable features to be considered

# scale data; find variable features, run pca and run umap
if(run) seurat.merged.processed <- process_obj(seurat.filtered.merged,ndim = ndim, nfeat = nfeat, find_variable_features = T) 
DimPlot(seurat.merged.processed,,split.by = "orig.ident")
```

# 4926 at 20%-90%
# 18.886 features 
```{r}
seurat.merged.processed
```

## run singleR
```{r}
if(run){
  ref.se <- celldex::ImmGenData()
  param <- BiocParallel::MulticoreParam(workers = 32)
  
  ## broad 
  result.singler.main = SingleR(test = GetAssayData(seurat.merged.processed), ref = ref.se, labels = ref.se$label.main, BPPARAM=param)
  seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = result.singler.main$pruned.labels,col.name = "SingleR.main")

  if(F){  
  # finer (ontology based)
    result.singler = SingleR(test = GetAssayData(seurat.merged.processed), ref = ref.se,labels = ref.se$label.ont, BPPARAM=param)
    # get term names
    cl <- ontoProc::getCellOnto()
    terms <- result.singler$pruned.labels%>%unique()
    getTerms <- ontoProc::label_TAG(ontology = cl,terms)
    seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = result.singler$pruned.labels,col.name = "SingleR.tag")
    seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = unname(getTerms[result.singler$pruned.labels]),col.name = "SingleR.term")
  }
}

```

How many LinB cells do we filtered out? 
```{r}
(seurat.merged.processed%>%subset(SingleR.main == "B cells"))$orig.ident%>%table()
```


# visualize SingleR classification

```{r,fig.height= 6,fig.width= 6}
Fig1a = DimPlot(seurat.merged.processed,group.by = "SingleR.main",label =T,repel =T,label.size = 3) + theme(aspect.ratio = 1) + labs(tag = "A")
ggsave(paste0(plot_folder,"/fig1a_singleR.pdf"),plot = Fig1a,width = 7, height = 7)
ggsave(paste0(plot_folder,"/fig1a_singleR.png"),plot = Fig1a,width = 7, height = 7)
ggsave(paste0(plot_folder,"/fig1a_singleR.tiff"),plot = Fig1a,width = 7, height = 7)
Fig1a

```

# Clustering whole assay
```{r}
if(run){
  set.seed(1234)
  seurat.merged.processed <- FindNeighbors(seurat.merged.processed, dims = 1:20)
  seurat.merged.processed <- FindClusters(seurat.merged.processed, resolution = 0.7)
}
```


```{r,fig.height= 5,fig.width= 5}
Fig1b  <- DimPlot(seurat.merged.processed,group.by = "seurat_clusters",label = T,label.size = 3) + theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + labs(tag = "B")
ggsave(paste0(plot_folder,"/Fig1b.pdf"), plot = Fig1b,width = 7, height = 7)
ggsave(paste0(plot_folder,"/Fig1b.png"),plot = Fig1b,width = 7, height = 7)
ggsave(paste0(plot_folder,"/Fig1b.tiff"),plot = Fig1b,width = 7, height = 7)
bcell_clusters = c("0","2","10","17")
Fig1b

```


```{r,fig.height= 4,fig.width= 12}

SFig1a <- DimPlot(seurat.merged.processed%>%subset(SingleR.main == "B cells"), group.by = "SingleR.main",label =T,repel =T,label.size = 3) +
  theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + ggtitle("SingleR B cells") + labs(tag = "A")
SFig1a

SFig1b <- DimPlot(seurat.merged.processed%>%subset(seurat_clusters %in% bcell_clusters),group.by = "seurat_clusters",label = T,label.size = 3) + 
  NoLegend() +  theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + ggtitle("Only BCell clusters (0,2,10,17)") + labs(tag = "B")

notBcell = setdiff(seurat.merged.processed$seurat_clusters%>%unique() ,bcell_clusters) 
SFig1c <- DimPlot(seurat.merged.processed%>%subset(seurat_clusters %in% notBcell),group.by = "seurat_clusters",label = T,label.size = 3)  + 
  NoLegend() +  xlim(-20, 15) + ylim(-12, 17) + ggtitle("Removing BCell clusters") + labs(tag = "C")

# By using qmin= 0.2 and qmax = 0.9 improves the dataset quality. SFig1d is no longer necesary.
#SFig1d <- DimPlot(seurat.merged.processed%>%subset(seurat_clusters %in% "7"),group.by = "seurat_clusters",label = T,label.size = 3) + 
#  NoLegend() +  theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + ggtitle("Mac-Mono Cluster (7)") + labs(tag = "D")



SFig1 <- SFig1a + SFig1b + SFig1c #+ SFig1d

ggsave(paste0(plot_folder,"/SFig1_singleR.pdf"),plot = SFig1,width = 15, height = 4.5)
ggsave(paste0(plot_folder,"/SFig1_singleR.png"),plot = SFig1,width = 15, height = 4.5)
ggsave(paste0(plot_folder,"/SFig1_singleR.tiff"),plot = SFig1,width = 15, height = 4.5)
SFig1

```



# scGate
The idea is to filter out some impure cells/doublets that could be passed thorough SingleR filtering model
```{r, fig.height= 4, fig.width= 5}
panBcell <- scGate::gating_model(name = "cd79", signature = "Cd79a")
panBcell <- scGate::gating_model(panBcell,name = "cd19", signature = "Cd19")
panBcell <- scGate::gating_model(panBcell, name = "cd20", signature = "Ms4a1")
panBcell <- scGate::gating_model(panBcell, name = "plasmaticas_residentes", signature = "Sdc1;Igha")
panBcell <- scGate::gating_model(panBcell, name = "Tcell", signature = "Cd3e;Cd3d;Cd4;Cd8a", negative = T)
panBcell <- scGate::gating_model(panBcell, name = "myeloid", signature = "Spi1", negative = T)
panBcell <- scGate::gating_model(panBcell, name = "MoMacDC", signature = "Lyz1;Lyz2;Csf1r;Msr1;Mafb;Cd300e", negative = T)
panBcell <- scGate::gating_model(panBcell, name = "Macrophage", signature = "Cd40;Cd68;Cd14;Fcgr1a", negative = T)
panBcell <- scGate::gating_model(panBcell, name = "Neutrophils", signature = "Cxcl2;Cd44;S100a9;Il1rn", negative = T)

```

```{r, fig.height= 4.5,fig.width= 5}
set.seed(1234)

seurat.merged.processed <- scGate(data = seurat.merged.processed, model = panBcell, output.col.name = "panBcell")

Fig1c <- DimPlot(seurat.merged.processed, cols = c(list(Impure = "gray", Pure = "green")), group.by = "panBcell" ) + 
  theme(aspect.ratio = 1) + ggtitle("scGate: panBcell model") + theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + labs(tag = "C")
seurat.merged.processed$panBcell%>%table()

ggsave(paste0(plot_folder,"/Fig1c.pdf"),plot = Fig1c,width = 7, height = 7)
ggsave(paste0(plot_folder,"/Fig1c.png"),plot = Fig1c,width = 7, height = 7)
ggsave(paste0(plot_folder,"/Fig1c.tiff"),plot = Fig1c,width = 7, height = 7)

Fig1c

```



## Filtering Bcells
```{r}
seurat.merged.processed$SinlgeR_Seurat_scGate = (seurat.merged.processed$SingleR.main == "B cells") & (seurat.merged.processed$seurat_clusters %in% bcell_clusters) & (seurat.merged.processed$panBcell == "Pure") 

seurat.merged.processed$SinlgeR_Seurat_scGate%>%table()
(seurat.merged.processed%>%subset(SinlgeR_Seurat_scGate == TRUE))$orig.ident%>%table()  

seurat.pan.Bcells <- seurat.merged.processed %>%subset(SinlgeR_Seurat_scGate == TRUE)
seurat.pan.Bcells
```



Manually remove outliers
```{r}

# esto estaba seteado en TRUE cuando usamos qmin = 0.1, qmax = 0.95 y hubo que sacar dos outliers a mano
if(F){
  SFig2 <- DimPlot(seurat.pan.Bcells,group.by = "SinlgeR_Seurat_scGate",label.size = 3) + theme(aspect.ratio = 1) +  xlim(-20, 15) + ylim(-12, 17)  + NoLegend() + ggtitle("SingleR + SeuratClusters + scGate")
  
  aux = DimPlot(seurat.pan.Bcells)
  seurat.pan.Bcells$outliers_umap <- (aux$data$UMAP_1 < -5 | aux$data$UMAP_2 < -1)
  seurat.pan.Bcells <- seurat.pan.Bcells%>%subset(outliers_umap == FALSE)
  seurat.pan.Bcells$orig.ident%>%table()  
  
  Fig1d <- DimPlot(seurat.pan.Bcells,group.by = "SinlgeR_Seurat_scGate",label.size = 3) + theme(aspect.ratio = 1) +  xlim(-20, 15) + ylim(-12, 17)  + NoLegend() + ggtitle("lin B cells")  + labs(tag = "D")
  

  ggsave(paste0(plot_folder,"/SFig2.pdf"),plot = SFig2,width = 7, height = 7)
  ggsave(paste0(plot_folder,"/SFig2.png"),plot = SFig2,width = 7, height = 7)
  ggsave(paste0(plot_folder,"/SFig2.tiff"),plot = SFig2,width = 7, height = 7)

  ggsave(paste0(plot_folder,"/Fig1d.pdf"),plot = Fig1d,width = 7, height = 7)
  ggsave(paste0(plot_folder,"/Fig1d.png"),plot = Fig1d,width = 7, height = 7)
  ggsave(paste0(plot_folder,"/Fig1d.tiff"),plot = Fig1d,width = 7, height = 7)

  SFig2 + Fig1d
  
}else{
  Fig1d <- DimPlot(seurat.pan.Bcells,group.by = "SinlgeR_Seurat_scGate",label.size = 3) + theme(aspect.ratio = 1) +  xlim(-20, 15) + ylim(-12, 17)  + NoLegend() + ggtitle("lin B cells")  + labs(tag = "D")

  ggsave(paste0(plot_folder,"/Fig1d.pdf"),plot = Fig1d,width = 7, height = 7)
  ggsave(paste0(plot_folder,"/Fig1d.png"),plot = Fig1d,width = 7, height = 7)
  ggsave(paste0(plot_folder,"/Fig1d.tiff"),plot = Fig1d,width = 7, height = 7)
  Fig1d
}  
```

```{r, fig.height= 10,fig.width= 12}
Fig1 <-  Fig1a + Fig1b + Fig1c + Fig1d
ggsave(paste0(plot_folder,"/Fig1.pdf"),plot = Fig1,width = 12, height = 10)
ggsave(paste0(plot_folder,"/Fig1.png"),plot = Fig1,width = 12, height = 10)
ggsave(paste0(plot_folder,"/Fig1.tiff"),plot = Fig1,width = 12, height = 10)
Fig1

```


## Save filtered dataset
```{r}
seurat.pan.Bcells
if(run){
  saveRDS(seurat.pan.Bcells,file = panBcellFile)
}
```

## Split WT and IM datasets and analize them
# Load Bcell-filtered dataset
```{r,warning=F, collapse =T,message =F}
ndim = 20; seed = 31416
sct_nFeatures <- 800

data <- readRDS(panBcellFile)
data%>%dim
data$condition <- ifelse(data$orig.ident == "DS-017-WT","wt","im") 
data.bkp <- data
data.wt <- data%>%subset(orig.ident == "DS-017-WT")
data.im <- data%>%subset(orig.ident != "DS-017-WT")
```

Visualize datasets by condition
```{r,fig.width=6,fig.height=3,warning=F, collapse =T,message =F}
fig2 <- DimPlot(data,group.by = "orig.ident",split.by = "orig.ident")
ggsave(paste0(plot_folder,"/Fig2.pdf"),plot = fig2,width = 12, height = 4)
ggsave(paste0(plot_folder,"/Fig2.png"),plot = fig2,width = 12, height = 4)
ggsave(paste0(plot_folder,"/Fig2.tiff"),plot = fig2,width = 12, height = 4)
fig2
```

# Preprocessing

**We will preprocess separately WT and IM subsets**

## Cell cycle scoring and regression 

The idea here is to mitigate the effects of cell cycle heterogeneity by calculating cell cycle phase scores based on canonical markers, and regressing these out of the data during pre-processing (see [SeuratVignette cell cycle](https://satijalab.org/seurat/articles/cell_cycle_vignette.html) for details)

```{r,warning=F, collapse =T,message =F, results='hide'}
# extract cell cycle genes from Seurat 
if(F){
  data.wt <- process_obj(data.wt,ndim = ndim)
  data.im <- process_obj(data.im,ndim = ndim)
}
s.genes <- cc.genes$s.genes%>%hs.2.mm(.)
g2m.genes <- cc.genes$g2m.genes%>%hs.2.mm(.)
```


```{r,warning=F, collapse =T,message =F, results='hide'}
# WT
set.seed(seed)
data.wt <- CellCycleScoring(data.wt, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
data.wt$CC.Difference <- data.wt$S.Score - data.wt$G2M.Score

data.wt <- SCTransform(data.wt, method = "glmGamPoi", vars.to.regress = c("percent.mito","CC.Difference"), verbose = FALSE, variable.features.n = sct_nFeatures, return.only.var.genes = FALSE)

data.wt <- RunPCA(data.wt, features = data.wt@assays$SCT@var.features,
                ndims.print = 1:5, nfeatures.print = 5)
data.wt <- RunUMAP(data.wt, reduction = "pca", dims = 1:ndim, seed.use=seed, n.neighbors = 20)
```



```{r,warning=F, collapse =T,message =F,results='hide'}
# IM
set.seed(seed)

data.im <- CellCycleScoring(data.im, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
data.im$CC.Difference <- data.im$S.Score - data.im$G2M.Score

# work as usual (only variable genes)
data.im <- SCTransform(data.im, method = "glmGamPoi", vars.to.regress = c("percent.mito","CC.Difference"), verbose = FALSE,variable.features.n = sct_nFeatures, return.only.var.genes = FALSE)

data.im <- RunPCA(data.im, features = data.im@assays$SCT@var.features,
                ndims.print = 1:5, nfeatures.print = 5)
data.im <- RunUMAP(data.im, reduction = "pca", dims = 1:ndim, seed.use=seed, n.neighbors = 20)
```


# Unsupervised analysis
```{r,fig.height=4, fig.width= 10,warning=F, collapse =T,message =F, results='hide'}
set.seed(seed)
data.im <- FindNeighbors(data.im, dims = 1:20)
#data.im <- FindNeighbors(data.im,assay  = "SCT", dims = 1:20)
data.im <- FindClusters(data.im, resolution = 0.7)
data.im <- FindClusters(data.im, resolution = 0.6)
data.im <- FindClusters(data.im, resolution = 0.5)
data.im <- FindClusters(data.im, resolution = 0.4)
data.im <- FindClusters(data.im, resolution = 0.3)
data.im <- FindClusters(data.im, resolution = 0.35)


set.seed(seed)
data.wt <- FindNeighbors(data.wt, dims = 1:20)
#data.wt <- FindNeighbors(data.wt,assay  = "SCT", dims = 1:20)
#data.wt <- FindClusters(data.wt, resolution = 2)
data.wt <- FindClusters(data.wt, resolution = 1.5)
data.wt <- FindClusters(data.wt, resolution = 1)
data.wt <- FindClusters(data.wt, resolution = 0.9)
data.wt <- FindClusters(data.wt, resolution = 0.8)
data.wt <- FindClusters(data.wt, resolution = 0.75)

data.wt <- FindClusters(data.wt, resolution = 0.7)

data.wt <- FindClusters(data.wt, resolution = 0.6)
data.wt <- FindClusters(data.wt, resolution = 0.65)

data.wt <- FindClusters(data.wt, resolution = 0.5)
data.wt <- FindClusters(data.wt, resolution = 0.4)
data.wt <- FindClusters(data.wt, resolution = 0.3)
```


```{r,fig.height=4, fig.width= 10,warning=F, collapse =T,message =F, results='hide'}
Fig3a <- DimPlot(data.wt,group.by = "SCT_snn_res.0.8",label = T)
Fig3b <- DimPlot(data.im ,group.by = "SCT_snn_res.0.4",label = T)

#Fig3a <- DimPlot(data.wt,group.by = "SCT_snn_res.0.8",label = T)
#Fig3b <- DimPlot(data.im ,group.by = "SCT_snn_res.0.4",label = T)

Fig3 <- Fig3a + Fig3b
Fig3

ggsave(paste0(plot_folder,"/Fig3a_clustering_wt.pdf"),plot = Fig3a,width = 4.5, height = 4) 
ggsave(paste0(plot_folder,"/Fig3a_clustering_wt.png"),plot = Fig3a,width = 4.5, height = 4)
ggsave(paste0(plot_folder,"/Fig3a_clustering_wt.tiff"),plot = Fig3a,width = 4.5, height = 4)

ggsave(paste0(plot_folder,"/Fig3b_clustering_im.pdf"),plot = Fig3b,width = 4.5, height = 4)
ggsave(paste0(plot_folder,"/Fig3b_clustering_im.png"),plot = Fig3b,width = 4.5, height = 4)
ggsave(paste0(plot_folder,"/Fig3b_clustering_im.tiff"),plot = Fig3b,width = 4.5, height = 4)


ggsave(paste0(plot_folder,"/Fig3.pdf"),plot = Fig3,width = 9, height = 4)
ggsave(paste0(plot_folder,"/Fig3.png"),plot = Fig3,width = 9, height = 4)
ggsave(paste0(plot_folder,"/Fig3.tiff"),plot = Fig3,width = 9, height = 4)

#FeaturePlot(data.im,features = c("nCount_RNA","nFeature_RNA"))  ## Visualizo para ver que no hay sesgo en los counts รณ nFeatures

```


## SingleR Sobre cada uno de estos datasets
```{r}
ref.se <- celldex::ImmGenData()
param <- BiocParallel::MulticoreParam(workers = 32)

if(run){
  
  if(T){  
    #finer (ontology based)
    result.singler = SingleR(test = GetAssayData(data.im,assay = "SCT",slot = "data"), ref = ref.se,labels = ref.se$label.ont, BPPARAM=param)
    # get term names
    cl <- ontoProc::getCellOnto()
    terms <- result.singler$pruned.labels%>%unique()
    getTerms <- ontoProc::label_TAG(ontology = cl,terms)
    data.im <- AddMetaData(data.im, metadata = result.singler$pruned.labels,col.name = "SingleR.tag")
    data.im <- AddMetaData(data.im, metadata = unname(getTerms[result.singler$pruned.labels]),col.name = "SingleR.term")
  }

}

if(run){
  if(T){  
    #finer (ontology based)
    result.singler = SingleR(test = GetAssayData(data.wt,assay = "SCT",slot = "data"), ref = ref.se,labels = ref.se$label.ont, BPPARAM=param)
    # get term names
    cl <- ontoProc::getCellOnto()
    terms <- result.singler$pruned.labels%>%unique()
    getTerms <- ontoProc::label_TAG(ontology = cl,terms)
    data.wt <- AddMetaData(data.wt, metadata = result.singler$pruned.labels,col.name = "SingleR.tag")
    data.wt <- AddMetaData(data.wt, metadata = unname(getTerms[result.singler$pruned.labels]),col.name = "SingleR.term")
  }
  
}



```




## Clean singleR # by filtering out predicted categories with less than ~1% cells
```{r,fig.width= 10,fig.height=4}
#wt
test <- data.wt
a <- test$SingleR.term%>%table()
categories_ok <- names(a[a>10])
categories_ok

clean.classif <- test$SingleR.term
clean.classif[!clean.classif%in%categories_ok] <- NaN

test$SingleR.term <- clean.classif
SFig2a <- DimPlot(test%>%subset(SingleR.term != "NaN"),group.by = "SingleR.term")
a <- test@meta.data[,c("SCT_snn_res.0.8","SingleR.term")]


#im
test <- data.im
a <- test$SingleR.term%>%table()
categories_ok <- names(a[a>20])
categories_ok

clean.classif <- test$SingleR.term
clean.classif[!clean.classif%in%categories_ok] <- NaN

test$SingleR.term <- clean.classif
SFig2b <- DimPlot(test%>%subset(SingleR.term != "NaN"),group.by = "SingleR.term")
a <- test@meta.data[,c("SCT_snn_res.0.4","SingleR.term")]



SFigSR <- SFig2a + SFig2b
SFigSR
ggsave(paste0(plot_folder,"/SFigSR.pdf"),plot = SFigSR,width = 14, height = 7)
ggsave(paste0(plot_folder,"/SFigSR.png"),plot = SFigSR,width = 14, height = 7)
ggsave(paste0(plot_folder,"/SFigSR.tiff"),plot = SFigSR,width = 14, height = 7)

```


## Write cluster ids to disk
```{r}
write.table(data.im$SCT_snn_res.0.4,paste0(markers_folder,"/clusters_im_200vfeat.tsv"),sep = "\t",quote = F,col.names = F)
write.table(data.wt$SCT_snn_res.0.8,paste0(markers_folder,"/clusters_wt_200vfeat.tsv"),sep = "\t",quote = F,col.names = F)
```


```{r}
saveRDS(data.wt,file = paste0(dataset_folder,"/data.wt.rds"))
saveRDS(data.im,file = paste0(dataset_folder,"/data.im.rds"))
```

```{r}
if(!run){
data.wt <- readRDS(paste0(dataset_folder,"/data.wt.rds"))
data.im <- readRDS(paste0(dataset_folder,"/data.im.rds"))
}
```


## #############################
# Marker analysis

```{r,fig.height= 13,fig.width=7,warning=F, collapse =T,message =F}
Idents(data.wt) <- "SCT_snn_res.0.8"

discrim.wt <- FindAllMarkers(data.wt, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.wt <- discrim.wt%>%subset(p_val_adj < 0.05)
gene.wt <- discrim.wt$gene
Fig4 <- DoHeatmap(data.wt,features = gene.wt)

Idents(data.im) <- "SCT_snn_res.0.4"
discrim.im <- FindAllMarkers(data.im,assay = "SCT", only.pos = TRUE, min.pct = 0.25,logfc.threshold = 0.15)
discrim.im.uncorrected <- discrim.im
discrim.im <- discrim.im%>%subset(p_val_adj < 0.05)

gene.im <- discrim.im$gene

Fig5a <- DoHeatmap(data.im,features = gene.im,size = 3,lines.width = 35)   # all clusters
```

## Remove cluster 5 in IM population (17 cells) (according to Carolina's inspection criteria)
```{r,fig.height= 13,fig.width=7,warning=F, collapse =T,message =F}
data.im.filt5 <- subset(data.im,SCT_snn_res.0.4!=5)
Idents(data.im.filt5) <- "SCT_snn_res.0.4"
discrim.im.filt5 <- FindAllMarkers(data.im.filt5,assay = "SCT", only.pos = TRUE, min.pct = 0.25,logfc.threshold = 0.15)
discrim.im.filt5.uncorrected <- discrim.im.filt5
discrim.im.filt5 <- discrim.im.filt5%>%subset(p_val_adj < 0.05)

gene.im.filt5 <- discrim.im.filt5$gene

Fig5b <- DoHeatmap(data.im.filt5,features = gene.im.filt5,size = 3,lines.width = 35)
```


```{r,fig.height= 7,fig.width=4,warning=F, collapse =T,message =F}
# show WT heatmap
Fig4
ggsave(Fig4,file = paste0(plot_folder,"/Fig4_heatmap_fdr005.wt.png"),height = 7,width = 4, dpi = 300)
ggsave(Fig4,file = paste0(plot_folder,"/Fig4_heatmap_fdr005.wt.tiff"),height = 7,width = 4, dpi = 300)
ggsave(Fig4,file = paste0(plot_folder,"/Fig4_heatmap_fdr005.wt.pdf"),height = 7,width = 4,dpi = 300)
write.table(discrim.wt, file = paste0(markers_folder,"/markers_fdr005.wt.tsv"),sep = "\t",quote =F,row.names =F)

```


```{r,fig.height= 15,fig.width=9,warning=F, collapse =T,message =F}
# show IM heatmap
Fig5a 
Fig5b 
ggsave(Fig5a,file = paste0(plot_folder,"/Fig5a_heatmap_fdr005.im.png"),height = 15,width = 9,dpi = 300)
ggsave(Fig5a,file = paste0(plot_folder,"/Fig5a_heatmap_fdr005.im.tiff"),height = 15,width = 9,dpi = 300)
ggsave(Fig5a,file = paste0(plot_folder,"/Fig5a_heatmap_fdr005.im.pdf"),height = 15,width = 9,dpi = 300)

ggsave(Fig5b,file = paste0(plot_folder,"/Fig5b_heatmap_fdr005.im.png"),height = 15,width = 9,dpi = 300)
ggsave(Fig5b,file = paste0(plot_folder,"/Fig5b_heatmap_fdr005.im.tiff"),height = 15,width = 9,dpi = 300)
ggsave(Fig5b,file = paste0(plot_folder,"/Figba_heatmap_fdr005.im.pdf"),height = 15,width = 9,dpi = 300)

write.table(discrim.im,file = paste0(markers_folder,"/markers_fdr005.im.tsv"),sep = "\t",quote =F,row.names =F)
```

```{r,eval = F}
# 
discrim.wt.down <- FindAllMarkers(data.wt, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.wt.down <- discrim.wt.down%>%subset(p_val_adj < 0.05)
write.table(discrim.wt.down,file =  paste0(markers_folder,"/markers_fdr005.wt_include_downReg.tsv"),sep = "\t",quote =F,row.names =F)


discrim.im.down <- FindAllMarkers(data.im, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.im.down <- discrim.im.down%>%subset(p_val_adj < 0.05)
write.table(discrim.im.down,file =  paste0(markers_folder,"/markers_fdr005.im_include_downReg.tsv"),sep = "\t",quote =F,row.names =F)


```



# DimPlot using Heatmap genes


WT dataset
```{r,fig.width= 10,fig.height= 5,warning=F, collapse =T,message =F}
#FeaturePlot(data.wt,features = discrim.wt$gene,ncol = 5)
```

IM dataset
```{r,fig.width= 20,fig.height=70,warning=F, collapse =T,message =F}
#FeaturePlot(data.im,features = intersect(discrim.im$gene,data.im%>%rownames()),ncol = 6)
```


# Explore predefined markers 

We will use here those genes previously defined in our molecular signature spreadsheet file

```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
features1 = c("Ighm","Ighd1-1","Igha","Ighg1","Cd19","Ms4a1","Fcer2a","Cd24a","Cd38","Pdcd1lg2","Nt5e","Entpd1","Cd80","Cd86","Cd93","Cd69","Itgal","Cd44","Cd28","Sdc1",	"Prdm1","Tnfrsf17","Sell","Igj")
features2 <- c("Fas","Bcl6","Cd40","Cxcr5","Cxcr4")
```

```{r,fig.height= 8,fig.width=8,warning=F, collapse =T,message =F}
h1 <- DoHeatmap(data.wt,features = features1)
h2 <- DoHeatmap(data.im,features = features1,size = 3,lines.width = 35)

h1
h2
```


WT dataset


```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = features1, ncol = 5)
```

IM dataset
```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = features1, ncol = 5)
```


WT dataset (new genes)
```{r,fig.height=7,fig.width=12,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = features2, ncol = 3)
```

IM dataset (new genes)
```{r,fig.height=7,fig.width=12,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = features2, ncol = 3)
```


# Explore markers of resident plasmatic cells (IgA+)
WT dataset
```{r,fig.width= 12,fig.height=8,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = c("Sdc1","Igha","Prdm1","Tnfrsf17","Cd28","Igj","Cd93"),ncol = 4)
```

IM dataset
```{r,fig.width= 13,fig.height=7,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = c("Sdc1","Igha","Prdm1","Tnfrsf17","Cd28","Igj","Cd93"),ncol = 4)
```



